\ProvidesFile{section_1_recap}[Section 1]
    
\newpage

\section{Introduction}

\begin{flushright}
    \emph{Machine learning is the most successful\\
    marketing campaign for rebranding linear algebra.}
    \end{flushright}

\subsection{What we will not do}
Create AGI. (Probably not.)

\subsection{What we will do}

Recap Linear Algebra 1. (I know that feel bro.)

Multiply matrices. Many matrices. By hand and on a computer. And then learn to see the geometric and physical meaning behind matrices and avoid unnecessary multiplications.

\subsection{Grading criteria}

\hspace{\parindent}30 \% --- In-class activities.

40 \% --- Graded home assignments.

30 \% --- In-class exams (midterm and final exams).

A bonus (1 \% of the total grade) will be given to those who bring truly funny and witty linear algebra memes that I haven't seen yet. (The degree of wit will be determined by my subjective sense of humor.)

\newpage
\section{Recap: vectors, matrices, determinants}
In this section we will recap the basic notions of linear algebra from Linear Algebra 1 that are used in this course without proofs.

Check your knowledge and see whether you can solve the following problems.

\begin{problem}
Consider the vector space of polynomials with real coefficients and zero constant term
$$
S = \{a_1 x + a_2 x^2 + \cdots + a_n x^n \mid a_i \in \R, n \in \N\}.
$$
Find a basis and the dimension of this vector space.
\end{problem}


\begin{problem}
Consider a set of all continuous functions on the interval $[0, 1]$, such that $f(x) \neq 0$ for all $x \in [0, 1]$. Is this set a vector space? If yes, find a basis and the dimension of this vector space. If no, explain why.
\end{problem}

% Kostrikin
\begin{problem}
    Let the vectors $e_1, e_2, e_3$ and $x$ be given by their coordinates in some basis:
    $$e_1=(2,1,-3), \quad e_2=(3,2,-5), \quad e_3=(1,-1,1), \quad x=(6,2,-7).$$
    Prove that $(e_1, e_2, e_3)$ is also a basis of the space, and find the coordinates of $x$ in this basis.    
\end{problem} 

\begin{problem}
    Prove that each of the two given systems of vectors $S$ and $S^{\prime}$ is a basis. Find the transition matrix $C$ from $S$ to $S^{\prime}$ ($e^{\prime} = e C$):
    $$
    \begin{aligned}
    & S=((1,2,1),(2,3,3),(3,8,2)) \\
    & S^{\prime}=((3,5,8),(5,14,13),(1,9,2))
    \end{aligned}
    $$
\end{problem}

\begin{problem}
    Prove that the given system of vectors is linearly independent, and complete it to a basis of the space of rows:
$$a_1=(2,2,7,-1), \quad a_2=(3,-1,2,4), \quad a_3=(1,1,3,1).$$
\end{problem}

\begin{problem}
    Find a basis and the dimension of the linear span of the following system of vectors:
$$a_1=(1,0,0,-1), \quad a_2=(2,1,1,0), \quad a_3=(1,1,1,1), \quad a_4=(1,2,3,4), \quad a_5=(0,1,2,3).$$
\end{problem}

\begin{problem}
 Given square matrices $A$ and $B$ of size $2025 \times 2025$ such that $\det(A) = 2$ and $\det(B) = 25$. Find $\det(BA)$.
\end{problem}

\begin{problem}
 Given a square matrix $A$ of size $2025 \times 2025$ with $\det(A) = 25$. Find $\det(A^{-1})$.
\end{problem}

\begin{problem}
 Find the rank of the matrix:
$$
A =\begin{pmatrix}
    1 & 1 & 3 & 4 & 5 \\
    0 & 0 & -1 & 2 & -1 \\
    0 & 0 & 0 & 14 & 15 \\
    0 & 0 & 0 & 28 & 30 \\
\end{pmatrix}.
$$
\end{problem}

\begin{problem}
     Find the dimension of the solution space of the homogeneous system $A x = 0$ with matrix:
    $$
    A =\begin{pmatrix}
    1 & 1 & 0 & 4 & 5 \\
    0 & 2 & 5 & 2 & -1 \\
    0 & 0 & 0 & 14 & 15 \\
\end{pmatrix}.
    $$
\end{problem}

% answer: -8
\begin{problem}
    Find the determinant of the following matrix:
$$
\begin{pmatrix}
    1 & 2 & 3 \\
    5 & 1 & 4 \\
    3 & 2 & 5 \\
\end{pmatrix}.
$$
\end{problem}

\begin{comment}
% answer: 0
\begin{problem}
    Find determinant of the following matrix:
    $$
    \begin{pmatrix}
    1 & 2 & 3 \\
    4 & 5 & 6 \\
    7 & 8 & 9 \\
    \end{pmatrix}.
    $$
    \end{problem}
\end{comment}

\begin{problem}
Let $A$ be a square $2 \times 2$ matrix with $\det(A) = 2$, $B$ be a square $5 \times 5$ matrix with $\det(B) = 5$, $C$ be a rectangular $2 \times 5$ matrix and $O$ be a zero matrix of size $5 \times 2$.
Find 
$$
\det\begin{pmatrix}
    A & C \\
    O & B \\
\end{pmatrix}.
$$
\end{problem}

\begin{problem}
 Find the determinant of the following $10 \times 10$ matrix:
$$
\begin{pmatrix}
    10 & 0 & 0 & \dots & 0 \\
    9 & 10 & 0 & \dots & 0 \\
    8 & 9 & 10 & \dots & 0 \\
    \vdots & \vdots & \vdots & \ddots & \vdots \\
    1 & 2 & 3 & \dots & 10 \\
\end{pmatrix}.
$$
\end{problem}

\begin{problem}
    Find the inverse of the following matrix:
$$
\begin{pmatrix}
    1 & 2 & 0 & 0 & 0 & 0 \\
    0 & 1 & 0 & 0 & 0 & 0 \\
    0 & 0 & 0 & 1 & 0 & 0 \\
    0 & 0 & 1 & 0 & 0 & 0 \\
    0 & 0 & 0 & 0 & 5 & 0 \\
    0 & 0 & 0 & 0 & 0 & 10
\end{pmatrix}.
$$
\end{problem}

\begin{comment}
\begin{problem}
    Find the inverse of the following matrix:
$$
\begin{pmatrix}
    10 & 0 & 0 & 0 & 0 & 0 \\
    0 & 10 & 0 & 0 & 0 & 0 \\
    0 & 0 & 1 & 0 & 0 & 0 \\
    0 & 0 & -2 & 1 & 0 & 0 \\
    0 & 0 & 0 & 0 & 0 & -1 \\
    0 & 0 & 0 & 0 & -1 & 0
\end{pmatrix}.
$$
\end{problem}
\end{comment}

\begin{problem}
Does the set of all invertible $5 \times 5$ matrices with real entries form a vector space? If yes, find a basis and the dimension of this vector space. If no, explain why.
\end{problem}

\begin{problem}
Does the set of all $5 \times 5$ matrices with real entries and zero trace ($a_{11} + a_{22} + a_{33} + a_{44} + a_{55} = 0$) form a vector space? If yes, find a basis and the dimension of this vector space. If no, explain why.
\end{problem}

\begin{problem}
Let $v$ be a vector with coordinates $(v_1, v_2, v_3)$ in the basis $e$ (where $e$ consists of 3 vectors), and let $C$ be the change-of-basis matrix ($e^{\prime} = eC$). Find the coordinates of $v$ in the basis $e^{\prime}$.  
\end{problem}